---
output: 
  stevetemplates::article:
    fig_caption: true
title: "Covariate Selection - Projection Prediction"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
# spacing: double
endnote: no
---

# Data Preprocessing

```{r, cache = F, include = F}
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
```

```{r, cache = F, include = F}
library(tidyverse)
library(brms)
library(projpred)
library(bayesplot)

# retrieve # of cores
ncores <- parallel::detectCores()

# for output clarity
options(scipen = 999)
```

```{r}
# load data
apes1 <- read_csv("../data/laac_data_trial.csv")
apes2 <- read_csv("../data/laac_data_task.csv")

fn0 <- function(x, ...) {
  # helper function
  # sum over correct choice variable (code)
  to_return = tibble(cogn = sum(x$code))
  return(to_return)
}

code_sum <- apes1 %>%
  # contains summed code variable [for each task, time point, session and subject]
  group_by(time_point, session, subject, task) %>%
  group_modify(fn0)

apes1_tmp <- apes1 %>%
  # helper for merging
  select(-c(date, trial_session, trial_time_point, code)) %>%
  unique(by = c("time_point", "session", "subject"))

apes1_new <-
  as_tibble(merge(apes1_tmp, code_sum, by = c("time_point", "session", "subject", "task"))) %>%
  mutate(across(c(subject, group, heat, test_day, le_present, dist_present, sex, rearing, observer), as_factor)) %>%
  mutate(observer = fct_relevel(observer, "no")) %>% 
  jtools::center(.,vars = c("sick_severity",
                            "le_mean",
                            "time_outdoors",
                            "age",
                            "time_in_leipzig")) %>%
  group_by(group, time_point) %>%
  mutate(rank_gmc = rank - mean(rank, na.rm = TRUE)) %>%
  ungroup() %>%
  arrange(time_point)

grp_size <- tibble(
  # number of apes for each species
  a_chimp = 20,
  b_chimp = 6,
  bonobo = 12,
  gorilla = 6,
  orangutan = 6
)

apes1_new <- apes1_new %>%
  # create rank variable depending on species
  group_by(group, time_point) %>%
  mutate(
    rel_rank = case_when(
      group == "a_chimp" ~ percent_rank(grp_size$a_chimp:1)[rank],
      group == "b_chimp" ~ percent_rank(grp_size$b_chimp:1)[rank],
      group == "bonobo" ~ percent_rank(grp_size$bonobo:1)[rank],
      group == "gorilla" ~ percent_rank(grp_size$gorilla:1)[rank],
      group == "orangutan" ~ percent_rank(grp_size$orangutan:1)[rank]
    )
  ) %>%
  ungroup()

apes1_new <- apes1_new %>%
  # create coding for heat variable
  mutate(heat_mod = case_when(
    sex == "f" & heat == "yes" ~ "_f_fheat",
    sex == "m" & heat == "yes" ~ "_m_fheat",
    sex == "f" & heat == "no" ~ "_f_noheat",
    sex == "m" & heat == "no" ~ "_m_noheat"),
    heat_mod = as_factor(heat_mod)
  ) %>% 
  mutate(heat_mod = fct_relevel(heat_mod, "_f_noheat"))

apes1_new <- apes1_new %>% 
  select(-heat, -heat_mod)

apes1_new <- apes1_new %>% 
  # recode rearing categories: hand -> unknown
  mutate(rearing = fct_recode(rearing, "hand" = "unknown"))

t_cau <- filter(apes1_new, task == "causality")
t_inf <- filter(apes1_new, task == "inference")
t_quant <- filter(apes1_new, task == "quantity")
t_gaze <- filter(apes1_new, task == "gaze_following")

t_gaze <- t_gaze %>%
  # create dummy variable indicating if session 1 or 2
  group_by(time_point, session) %>% 
  mutate(tp_mod = cur_group_id()) %>%
  ungroup() %>% 
  mutate(day2 = case_when(session == 1 ~ "no",
                          session == 2 ~ "yes"),
         day2 = factor(day2)) %>%
  select(tp_mod, day2, everything())

t_gaze <- t_gaze %>%
  # remove duplicates created by day2
  group_by(subject) %>%
  filter(!duplicated(tp_mod)) %>%   
  ungroup() 
```

# Covariate Selection

```{r}
# covariate needed for projection prediction
# placed here to compare with formula
all_fixed_effects <- c("sick_severity",       
                       "test_day", "test_tp", 
                       "rel_rank",
                       "observer", 
                       "age", "time_in_leipzig", 
                       "sex", 
                       "rearing", 
                       "le_mean", 
                       "dist_mean",
                       "time_outdoors", 
                       "sociality")

fm <- formula(cogn ~ sick_severity +                    
                test_day + test_tp +                
                rel_rank + # rank_gmc +
                observer + 
                age + time_in_leipzig +
                sex + 
                rearing +
                le_mean + # le_max + # le_present +         
                dist_mean + # dist_max + # + dist_present +
                time_outdoors +
                sociality + # sociality_total
                # heat_mod + # heat +
                (1|subject)                         
              ) 

fm_gaze <- update(fm, . ~ . +day2)
```

## Reference Model: 2-level Multilevel Model (random intercepts only)

```{r}
m_cau_2l <- brm(fm, data = t_cau, 
                warmup = 1e3, iter = 3e3, cores = ncores, chains = 2, 
                seed = 2021,
                save_pars = save_pars(all = TRUE)
                )
m_inf_2l <- brm(fm, data = t_inf, 
                warmup = 1e3, iter = 3e3, cores = ncores, chains = 2, 
                seed = 2021,
                save_pars = save_pars(all = TRUE)
                )
m_quant_2l <- brm(fm, data = t_quant, 
                  warmup = 1e3, iter = 3e3, cores = ncores, chains = 2, 
                  seed = 2021,
                  save_pars = save_pars(all = TRUE)
                  )
m_gaze_2l <- brm(fm_gaze, data = t_gaze, 
                 warmup = 1e3, iter = 3e3, cores = ncores, chains = 2, 
                 seed = 2021,
                 save_pars = save_pars(all = TRUE)
                 )
```

```{r}
summary(m_cau_2l)
summary(m_inf_2l)
summary(m_quant_2l)
summary(m_gaze_2l)
```

## Predictive Projection

```{r}
# delay random intercept to last place so that it doesn't soak up all the variance
s_terms <- c("1", all_fixed_effects, 
             paste0(paste(all_fixed_effects, collapse = " + "), " + (1 | subject)")) 
s_terms_gaze <- c("1", c(all_fixed_effects, "day2"), 
                  paste0(paste(c(all_fixed_effects, "day2"), collapse = " + "), " + (1 | subject)"))
```

```{r}
refM_cau <- get_refmodel(m_cau_2l)
refM_inf <- get_refmodel(m_inf_2l)
refM_quant <- get_refmodel(m_quant_2l)
refM_gaze <- get_refmodel(m_gaze_2l)
```

```{r}
vs_cau <- varsel(refM_cau, search_terms = s_terms, ndraws_pred = 400)
summary(vs_cau); plot(vs_cau, stats = c('elpd', 'rmse'))
suggest_size(vs_cau)

randint_ind_vscau <- length(solution_terms(vs_cau))
relevant_cov_vscau <- c(1, 2, 3, 4, 5, randint_ind_vscau)

proj_cau <- project(vs_cau, solution_terms = relevant_cov_vscau)
mcmc_areas(as.matrix(proj_cau), pars = solution_terms(vs_cau)[relevant_cov_vscau])
```

```{r}
vs_inf <- varsel(refM_inf, search_terms = s_terms)
summary(vs_inf); plot(vs_inf, stats = c('elpd', 'rmse'))

randint_ind_vsinf <- length(solution_terms(vs_inf))
relevant_cov_vsinf <- c(1, 2, 3, 4, 5, randint_ind_vsinf)

proj_inf <- project(vs_inf, solution_terms = relevant_cov_vsinf)
mcmc_areas(as.matrix(proj_inf), pars = solution_terms(vs_inf)[relevant_cov_vsinf])
```

```{r}
vs_quant <- varsel(refM_quant, search_terms = s_terms)
summary(vs_quant); plot(vs_quant, stats = c('elpd', 'rmse'))

randint_ind_vsquant <- length(solution_terms(vs_quant))
relevant_cov_vsquant <- c(1, 2, 3, 4, 5, randint_ind_vsquant)

proj_quant <- project(vs_quant, solution_terms = relevant_cov_vsquant)
mcmc_areas(as.matrix(proj_quant), pars = solution_terms(vs_quant)[relevant_cov_vsquant])
```

```{r}
vs_gaze <- varsel(refM_gaze, search_terms = s_terms_gaze)
summary(vs_gaze); plot(vs_gaze, stats = c('elpd', 'rmse'))

randint_ind_vsgaze <- length(solution_terms(vs_gaze))
relevant_cov_vsgaze <- c(1, 2, 3, 4, 5, randint_ind_vsgaze)

proj_gaze <- project(vs_gaze, solution_terms = relevant_cov_vsgaze)
mcmc_areas(as.matrix(proj_gaze), pars = solution_terms(vs_gaze)[relevant_cov_vsgaze])
```

## Covariate Selection (cross validated)

```{r}
cvs_cau <- cv_varsel(refM_cau, search_terms = s_terms, cv_method = "loo", method = "forward")
summary(cvs_cau); plot(cvs_cau, stats = c('elpd', 'rmse'))
proj_cau_cv <- project(cvs_cau, solution_terms = c(1, 2, 3, 14))
mcmc_areas(as.matrix(proj_cau_cv), pars = solution_terms(cvs_cau)[c(1, 2, 3, 14)])
```
**relevant** covariates: (1 | subject), sick_severity, dist_mean, time_in_leipzig

```{r}
cvs_inf <- cv_varsel(refM_inf, search_terms = s_terms, cv_method = "loo", method = "forward")
summary(cvs_inf); plot(cvs_inf, stats = c('elpd', 'rmse'))
proj_inf_cv <- project(cvs_inf, solution_terms = c(1, 2, 14))
mcmc_areas(as.matrix(proj_inf_cv), pars = solution_terms(cvs_inf)[c(1, 2, 14)])
```
**relevant** covariates: (1 | subject), time_in_leipzig, age

```{r}
cvs_quant <- cv_varsel(refM_quant, search_terms = s_terms, cv_method = "loo", method = "forward")
summary(cvs_quant); plot(cvs_quant, stats = c('elpd', 'rmse'))
proj_quant_cv <- project(cvs_quant, solution_terms = c(1, 2, 3, 4, 5, 6, 7, 14))
mcmc_areas(as.matrix(proj_quant_cv), pars = solution_terms(cvs_quant)[c(1, 2, 3, 4, 5, 6, 7, 14)])
```
**relevant** covariates: (1 | subject), time_in_leipzig, rearing, rel_rank, observer, test_tp, sex, sick_severity

```{r}
cvs_gaze <- cv_varsel(refM_gaze, search_terms = s_terms_gaze, cv_method = "loo", method = "forward")
summary(cvs_gaze); plot(cvs_gaze, stats = c('elpd', 'rmse'))
proj_gaze_cv <- project(cvs_gaze, solution_terms = c(1, 2, 3, 4, 5, 6, 7, 15))
mcmc_areas(as.matrix(proj_gaze_cv), pars = solution_terms(cvs_gaze)[c(1, 2, 3, 4, 5, 6, 7, 15)])
```
**relevant** covariates: (1 | subject), time_outdoors, observer, sick_severity, sex, sociality, rearing, age

Choosing the relevant covariates is based on inspection elpd/rmse plots.

## pareto $k$ values too high

```{r}
#library(loo)
#loo(m_cau_2l)
#loo(m_inf_2l)
#loo(m_quant_2l)
#loo(m_gaze_2l)
```

Pareto $k$ values are ok.



